A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2018-12-02 02_Materials_and_Methods.Rmd
problems of Radon-B

The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible: **doi**.

The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013]. It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC). Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B (as of December 2018) is a collection of more than 11000 published dates from around 2800 sites from all over europe. Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the the type and cultural attribution of the sampled context. The variable *Feature type* systematically distinguishes for example between settlements, rockshelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, etc.. This information allows to accurately and precisely determine a moment in space and time, when a grave of a certain type was created. In theory the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe. Unfortunately the Radon-B dataset has some data quality issues: **TODO**. The data representativity and semantic consequences will be discussed below.

The following major steps of Data preparation were applied. Please see the code for the details of implementation. The value in brackets indicates the number of dates still available after the respective filtering operation.

>>>>

The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible: **doi**. A more in-depth description and discussion of the data preparation can be found in the master thesis this article is based on **TODO**.

The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013]. It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC). Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B is a collection of more than 11000 published dates from around 2800 sites from all over europe (as of December 2018). Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the the type and cultural attribution of the sampled context. The variable *Feature type* systematically distinguishes for example between settlements, rockshelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, etc.. This information allows to accurately and precisely determine a moment in space and time, when a grave of a certain type was created. In theory the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe. Unfortunately the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no supervised or well thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries. These issues though do not inhibit the general data analysis strategy of this paper.

To create a tidy and useful subset of Radon-B, the following major steps of data preparation were applied. Please see the code for details of implementation. The value in brackets indicates the number of dates still available after the respective filtering operation.
